% SPDX-FileCopyrightText: 2023 Iegor Riepin, Tom Brown
%
% SPDX-License-Identifier: CC-BY-4.0

% Introduction Structure

% 1.1 Opening: ICT sector energy consumption and emissions

Demand for digital services is growing rapidly. The number of internet users has doubled worldwide over the past decade, while traffic has grown 25-fold \cite{ieaDataCentresData2023}.
The significant energy consumption of Information and Communications Technology (\gls{ict}) infrastructure is becoming a concern from an ecological standpoint.
Global datacenter energy use is estimated to be between 1\% and 2\% of final electricity demand worldwide in 2020 \cite{myttonHowMuchEnergy2020, masanetRecalibratingGlobalData2020} and is likely to increase rapidly in the future \cite{andraeGlobalElectricityUsage2015}.
The Greenhouse Gas (\gls{ghg}) emissions footprint of the datacenters and data transmission networks was estimated to be 330~million tones (Mt) \co-eq in 2020, equivalent to 0.9\% of global energy-related emissions \cite{ieaDataCentresData2023, malmodinICTSectorElectricity2023}.
For perspective, the national CO$_2$ emissions of the United Kingdom were around 331~Mt in 2021 \cite{UKnationalstats}.

% **datacenter special features as electricity consumers*

The growing demand for computing drives innovation and economies of scale. Large companies, such as Amazon, Google, and Microsoft are centralizing computing infrastructure into \enquote{hyperscale} datacenters.
Such large and highly efficient facilities consist of hundreds of server nodes that are managed collectively and geographically dispersed throughout the world \cite{ThereAre500}.

In their role as electricity consumers, hyperscale datacenters have unique characteristics when it comes to demand-side management.
Some computing jobs and associated power loads are \enquote{flexible}, i.e., the jobs are not time-critical and can be scheduled flexibly without impacting the overall quality of service.
Therefore, datacenter operators have the ability to shift a considerable fraction of computing workloads \textit{in~time} and \textit{in space} \cite{radovanovicIEEE2023}.
The temporal load shifting implies re-scheduling of flexible workloads to another time point, i.e., delaying computing job execution.
The spatial load shifting implies migration of flexible compute jobs and associated power loads between different physical datacenter locations.

The significant energy consumption of \gls{ict} infrastructure, and its anticipated growth, raise a natural question: \textit{How can datacenters make use of their unique spatial and temporal load-shifting capabilities to improve their economics and the environmental footprint?} The topic of this overarching question has been the subject of a growing body of literature over the past decade.

% Literature review
\textbf{Existing literature --} The work by Wierman et al.  \cite{wiermanOpportunitiesChallengesData2014} surveyed the forms, opportunities and challenges for data center demand response options, including aspects of temporal load shifting such as workload delaying or shedding and the geographical load balancing. Several papers discussed methods and policies needed to incentivise parcicipation of datacenters in the power grid's demand responce \cite{liuPricingDataCenter2014, zhouTruthfulEfficientIncentive2020}.  New advancements in this area include the application of quantum computing techniques for energy management in data centers, as demonstrated by Ajagekar and You \cite{ajagekarVariationalQuantumCircuit2024}, showcasing potential for significant reductions in power consumption and carbon emissions.

Numerous studies investigated ways to harness spatial or temporal load-shifting flexibility to achieve certain economical or environmental goals. Many studies have proposed workload scheduling algorithms for geographical load shifting to reduce datacenter electricity costs \cite{velascoElasticOperationsFederated2014, douCarbonAwareElectricityCost2017, heMinimizingOperationCost2021, raoDistributedCoordinationInternet2012, renCarbonAwareEnergyCapacity2012, dengEcoAwareOnlinePower2016}, or looked at the potential of datacenter load migration for mitigating renewable curtailment and \gls{ghg} emissions in the local electricity grid \cite{zhengMitigatingCurtailmentCarbon2020, mahmudDistributedFrameworkCarbon2016}, or aimed to increase the share of renewable energy in datacenter electricity consumption \cite{wangGreenawareVirtualMachine2015, kimDataCentersDispatchable2017, liuGeographicalLoadBalancing2011, kellyBalancingPowerSystems2016}.

Recently, several papers elaborated a mathematical problem that captures both spatial and temporal load-shifting flexibility. Zhang et al. \cite{zhangFlexibilityNetworksData2020} introduced the concept of virtual links to capture space-time load flexibility provided by geographically-distributed data centers in market clearing procedures. This work was followed by a paper in which the authors demonstrated a market clearing formulation that seeks to remunerate spatio-temporal load shifting for the flexibility service datacenters can offer to the power grid \cite{zhangRemuneratingSpaceTime2022}.

A great deal of attention in previous research exploring the signals for geographical load shifting has focused on information characterising the state of electricity grids where datacenters operate. These signals include average carbon emissions in the region of operation \cite{zhengMitigatingCurtailmentCarbon2020}, or locational marginal carbon emissions \cite{lindbergEnvironmentalPotentialHyperScale2021}, or locational electricity prices and price differences across datacenter locations \cite{raoMinimizingElectricityCost2010,tranHowGeoDistributedData2016, zhangRemuneratingSpaceTime2022}. This emerging stream of research work is now finding practical applications. For example, Google recently introduced a Carbon-Intelligent Compute Management system that minimizes carbon footprint and power infrastructure costs by shifting flexible workloads in time  across datacenter fleet as a function to the next day's carbon intensity forecasts \cite{radovanovicIEEE2023}. There are several companies providing market data and forecasts on carbon emissions and electricity prices, such as WattTime \cite{WattTime}, ElectricityMaps \cite{ElectricityMaps}, and others.

% 24/7 CFE**
\textbf{24/7 Carbon-Free Energy matching --} \gls{ict} companies invest significantly in renewable energy to reduce their environmental impact, reduce power price volatility, and enhance their brand image. The hyperscale datacenter industry in particular leads in corporate renewable energy acquisition through Power Purchase Agreements (\gls{ppa}s). With almost 50~GW contracted so far, Amazon, Microsoft, Meta, and Google are the four largest purchasers of corporate renewable energy \gls{ppa}s \cite{ieaDataCentresData2023}. Furthermore, some companies have pledged to reduce \textit{all} greenhouse gas emissions associated with their electricity use. These companies have set a \enquote{24/7 Carbon-Free Energy (\gls{cfe})} goal aiming to match electricity demand with clean energy supply on an \textit{hourly basis}. The 24/7 CFE targets were announced by Google and Microsoft for 2030 and Iron Mountain for 2040 \cite{google-247by2030, Microsoft-vision, IronMountainSustainability}. In the series of position papers, Google has outlined the roadmap to achieve 24/7 CFE and the methodologies to measure progress towards this goal \cite{google-methodologies,google-PolicyRoadmap}.

One notable feature of the 24/7 CFE goal is that it drives---and requires---a significant volume of energy procurement via \gls{ppa}s. Recent research showed that achieving 24/7 CFE with a 100\% target (i.e., achieving 0~gCO$_2$ emissions per kWh) implies that companies need to rely on their own procured resources,without any electricity procurement from the local grid \cite{riepinMeansCostsSystemlevel2023, riepin-zenodo-systemlevel247}. This is because the local electricity mix in most regions and at most times does not have a strictly zero carbon content and therefore cannot contribute to 24/7 CFE.

% The Problem**
\textbf{The problem --}
Companies aiming for 24/7 clean electricity have a unique challenge regarding load-shifting. Because they rely on their own portfolio of carbon-free energy generators instead of buying electricity from a local grid, grid signals such as average carbon emissions or locational electricity prices no longer play a significant role for informed load-shifting. This effect was illustrated with the help of energy system modelling in the previous work of the authours \cite{riepin-zenodo-systemlevel247,riepinMeansCostsSystemlevel2023}. In this context, the following question arise: \textit{In pursuit of achieving 24/7 Carbon-Free Energy objectives, what signals should datacenter operators focus on to facilitate informed and effective load shifting across space and time?}

% 6. **Relevance of this paper**
\textbf{Contribution --} This paper is the first to examine the role of spatio-temporal load shifting in the context of 24/7 carbon-free energy (CFE) matching. Using energy system modeling, we identify and investigate signals that are critical for effective load shifting across space and time to meet 24/7 CFE objectives. Specifically, we isolate three key signals and demonstrate that incorporating them into load-shifting and energy procurement strategies enables companies to achieve substantial gains in both resource efficiency and cost-effectiveness for clean computing.

For that, we develop a computer model that simulates a datacenter fleet with controlled degrees of flexibility in pursuit of 24/7 CFE objectives.
By incorporating datacenter fleet optimization into a European electricity system model as a mathematical problem, our analysis uncovered three individual signals for effective geographical and temporal load shifting: (a) varying quality of renewable energy resources (i.e. average capacity factors) across datacenter locations, (b) low correlation between wind power generation over long distances due to different weather conditions, and (c) lags in solar radiation peak due to Earth's rotation.
The results show that each of these signals plays a role in effective load shifting, although the \enquote{weight} of each signal in optimal load shaping strategies depends on the locations of datacenters and time of the year.
Furthermore, we demonstrate that load shifting decisions optimized considering these signals facilitate efficiency and affordability of the 24/7 CFE matching. When all three signals are taken into account, the electricity costs of a datacenter fleet operating at 0~kgCO$_2$/MWh can be reduced by 1.29$\pm$0.07~\euro/MWh for each percent point of flexible load by means of spatio-temporal load shifting.

% 7. **Prologue**

The remainder of the paper is structured as follows: \cref{sec:methods} introduces a mathematical model of clean energy procurement and gives a brief summary of the methodology, sources of model input data, and the experimental setup. \cref{sec:results} presents the results of the experiments and discusses the implications of the findings. \cref{sec:discussion} discusses the results in the context of the existing literature, practical recommendations for incorporating discussed signals into load shaping strategies and gives critical appraisal of the results. \cref{sec:conclusion} concludes the paper. All code necessary to reproduce the experiments is published under an open license alongside the paper. See \nameref{sec:code}.

